### 2020-04-07 projects/asparagus/doc/slides.tex
updated slides.tex in projects/asparagus/doc

\documentclass[aspectratio=1610, xcolor=dvipsnames,handout]{beamer} %aspectratio=169, 
\usetheme{Pittsburgh}
\usepackage{tikz}
\usepackage{pgfplots}
\usepackage{booktabs}
\usepackage[version=4]{mhchem}
\usepackage{animate}
\usepackage{biblatex}
\usepackage[absolute,overlay]{textpos}
\usepackage{fancyvrb}
\usepackage{xcolor}
\usepackage{multicol}
\usepackage{setspace}
\usepackage{siunitx}
\DeclareSIUnit\year{a}
\usepackage{qrcode}
\usepackage{eurosym}
\usepackage{xfrac}
\usepackage[export]{adjustbox}

% Thin fonts
\usepackage{cmbright}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{graphicx}
\usepackage{amsmath}

\pgfplotsset{compat = 1.17}
\usetikzlibrary{arrows.meta}
\usetikzlibrary{shapes.geometric, arrows}
\urlstyle{same}

%gets rid of bottom navigation bars
\setbeamertemplate{footline}[frame number]{}
%gets rid of bottom navigation symbols
\setbeamertemplate{navigation symbols}{}

% colors
\definecolor{bokugreen}{rgb}{0.302, 0.675, 0.149}
\setbeamercolor{title}{fg=bokugreen, bg=white}
\setbeamercolor{frametitle}{fg=bokugreen}
\setbeamercolor{itemize item}{fg=bokugreen}
\setbeamercolor{enumerate item}{fg=bokugreen}
\setbeamercolor{local structure}{fg=bokugreen}

\newcommand{\backupbegin}{
   \newcounter{framenumberappendix}
   \setcounter{framenumberappendix}{\value{framenumber}}
}
\newcommand{\backupend}{
   \addtocounter{framenumberappendix}{-\value{framenumber}}
   \addtocounter{framenumber}{\value{framenumberappendix}} 
}

% title page
\setbeamertemplate{title page}[default][colsep=-4bp,rounded=true]
\title[The cost of undisturbed landscapes]{The cost of undisturbed landscapes}
\subtitle{Towards designing energy systems from a social perspective}
\author{Sebastian Wehrle, Johannes Schmidt}
\institute{University of Natural Resources and Life Sciences, Vienna}
\date{May 7, 2020\\ \medskip EGU General Assembly 2020}

%\setbeamertemplate{section page}{
%	\begin{center}
%	{\usebeamerfont[fg]{section title} \insertsection}
%	\end{center}
%}
\AtBeginSection[]{
  \begin{frame}
  \vfill
%  \centering
%  \begin{beamercolorbox}[sep=8pt,center,shadow=false,rounded=true]{title}
\tableofcontents[currentsection]
%\begin{textblock*}{10cm}(7cm, 2.0cm) % {block width} (coords)
%\adjincludegraphics[height=6cm,trim={0 0 {.3\width} 0},clip]{./figures/pv}
%\end{textblock*}
 %   \usebeamerfont{title} \usebeamercolor{title} \insertsection\par%
%  \end{beamercolorbox}
  \vfill
  \end{frame}
}

\begin{document}
    %%%%Title Page
    {
        \usebackgroundtemplate{
	        \begin{picture}(320,325)
    	        \hspace{9.0cm}
	              \includegraphics[width=0.7\linewidth]{./logos/refuel_logo_with_text}
            \end{picture}
        }

        \begin{frame}[noframenumbering,plain]
          \maketitle
          \vspace{1.15cm}
          \includegraphics[height=0.7cm]{./logos/creative-commons-by}
          \hspace{4.0cm}\includegraphics[height=1.7cm]{./logos/boku-logo}\\
         \end{frame}
    }

    \section{Background \& Motivation}\label{sec:background}
    \begin{frame}
        \frametitle{State of the Art}
        Power system models used to
        \begin{itemize}
            \item demonstrate the feasibility of highly renewable energy systems
            \item determine the system cost-minimizing mix of renewable energy technologies
        \end{itemize}
        find significant capacities of renewable energy technologies (RET), largely wind power, need to be added.
        \medskip

        Researchers are increasingly aware that social and political restrictions on such large-scale RET expansions must be
        recognized.
        \begin{itemize}
            \item RET deployment in power system models often limited on the upside by results from analyses of land
            availability, e.g. by excluding areas for technical or legal reasons, or because they are constested by stakeholders
        \end{itemize}
        \medskip

        Yet, such `binary' analysis neglects the trade-offs faced by a social planner deciding on RET expansion.
    \end{frame}

    \begin{frame}
        \frametitle{Motivation}
        \begin{itemize}
            \item So far, analysis of \emph{system cost} minimizing energy systems
            \begin{itemize}
                \item Investment cost
                \item Operation and maintenance cost (including fuel and emission cost)
                \smallskip
                \item[$\rightarrow$] includes external effects of RET on power system, such as
                \begin{itemize}
                    \item Profile cost
                    \item Transmission expansion cost
                    \item Balancing cost
                \end{itemize}
            \end{itemize}
        \end{itemize}
        \medskip
        \begin{itemize}
            \item We need analysis of \textbf{\emph{social cost}} minimizing energy systems, including
            %\item We need to develop tools to minimize \textbf{social cost} of power systems, including
            \begin{itemize}
                \item Local external cost, such as visual impact on landscape, harm to wildlife, noise, glaring, \ldots
                \item Benefits of system transition, such as avoided cost of local air pollution
                \item[$\rightarrow$] includes external effects of RET beyond power system
            \end{itemize}
        \end{itemize}
    \end{frame}

    \begin{frame}
        \frametitle{Framework}

        \begin{textblock*}{5cm}
            (1.0cm,2cm)
            Consider a renewable energy technology that
            \begin{itemize}
                \item saves system cost (relative to its best alternative), but
                \item comes at social cost (e.g. due to local negative external effects)
            \end{itemize}
            \medskip

            A social planner should deploy this RET (instead of another one) up to the capacity at which
            \begin{itemize}
                \item system cost saving equals
                \item added social cost
            \end{itemize}
        \end{textblock*}

        \begin{textblock*}{8.0cm}
            (6.5cm,2cm)
            \begin{tikzpicture}
                \begin{axis}[
                    width=\textwidth,
                    clip=false,
                    axis lines=left,
                    ticks=none,
                    xlabel={capacity substituted},
                    ylabel={cost},
                    domain=0.5:1.75,
                    xmin=0.5, xmax=2,
                    ymin=0, ymax=4,
                    samples=200
                ]
                    \addplot[blue, thick] (x, 1.3*x^1.1+0.6);
                    \node at (axis cs: 1.925,3) [blue, align=left] {saved\\ system\\ cost};

                    \addplot[bokugreen, thick] (x, -2*x^0.8+4);
                    \node at (axis cs: 2.0,1) [bokugreen, align=left]{harm \\(from RET\\ externality)};
                \end{axis}
            \end{tikzpicture}
        \end{textblock*}
    \end{frame}

    \begin{frame}
        \frametitle{Approach}
        Focus on the social cost of power systems, including local externalities of \emph{wind power}
        \begin{itemize}
            \item Seems to be at center of societal debate currently
            \item Local externalities well researched (compared to externalities of open-space PV)
        \end{itemize}
        \medskip
        An assessment of socially optimal renewables expansion must be based on two pillars
        \begin{enumerate}
            \item \textbf{\textcolor{bokugreen}{The cost of undisturbed landscapes}}
            \begin{itemize}
                \item[--] energy system costs of substituting wind power with other RET
                \item[--] the \emph{opportunity cost} of wind power (versus its best alternative)
            \end{itemize}
            \item \textcolor{bokugreen}{The benefit of undisturbed landscapes}
            \begin{itemize}
                \item[--] avoided (local) negative external effects of wind turbines
                \item[--] avoided (local) external cost of wind power
            \end{itemize}
        \end{enumerate}
    \end{frame}

    \section{Method}\label{sec:method}

    \begin{frame}
        \frametitle{The cost of undisturbed landscapes}
        \framesubtitle{Estimating the opportunity cost of wind power}
        To determine the opportunity cost of substituting wind power with solar PV, we
        \begin{enumerate}
            \item derive the unrestricted system cost-minimizing deployment of wind and solar power \label{step1}
            \item restrict deployment of wind power by a small margin (so that the next best RET substitutes for wind power) and
            observe net system cost $c_{net}$ (calculated as total system cost including air pollution cost net of trade balance) \label{step2}
            \item repeat \ref{step2} till no wind power can be deployed
            \item the opportunity cost of wind power is approximated by the change in $c_{net}$ in response to a change in wind
            power capacity $w$ deployed.
            \begin{align*}
                OC_w = \frac{\Delta c_{net}}{\Delta w}
            \end{align*}
        \end{enumerate}
    \end{frame}

    \begin{frame}
        \frametitle{The case of Austria}
        \begin{itemize}
            \item Policy goal: \SI{100}{\percent}\footnote[frame]{excluding system services and industry own consumption.
            At current levels this equals 10\% of consumption, i.e. actual target is around 90\%.} of electricity demand from
            domestic renewable sources on annual balance by 2030
            \begin{itemize}
                \item[$\rightarrow$] Ideal frame for studying substitution / opportunity cost
            \end{itemize}
            \item necessarily turns Austria into a net exporter of electricity
            \item Official expectation: additional generation of \SI{27}{\tera\watt\hour} sufficient
        \end{itemize}
        \begin{table}
            \centering
            \begin{tabular}{| l | r | r | r | r | r |}
                \hline
                & 2018 & 2018\footnote[frame]{meteorological conditions as in 2016} & Policy & 2030 & 2030 \\
                & [GW] & [TWh] & [TWh] & [TWh] & [GW] \\ \hline \hline
                Solar PV & 1.44 & 1.23 & +11 & 12.23 & 14.27 \\ \hline
                Wind (onshore) & 3.05 & 6.14 & +10 & 16.14 & 8 \\ \hline
                Hydro (run-of-river) & 5.72 & 28.34 & +5 & 33.34 & 6.73 \\ \hline
                Biomass & $\sim 1$ & 4.78 & +1 & 5.78 & $<2$ \\ \hline
            \end{tabular}
        \end{table}
    \end{frame}

%\begin{frame}
%\frametitle{Austrian energy policy objectives}
%\framesubtitle{according to government programme 2020--2024}
%\begin{itemize}
%\item 100\%\footnote[frame]{excluding system services and industry own consumption.
%    At current levels this equals 10\% of consumption, i.e. actual target is around 90\%.} of electricity demand from
%    domestic renewable sources on annual balance by 2030 \pause
%\begin{itemize}
%\item[$\rightarrow$] necessarily turns Austria into a net exporter of electricity \pause
%\end{itemize}
%\item additional annual electricity generation of \SI{27}{\tera\watt\hour} is expected to suffice \pause
%\item technology-specific additions:
%\end{itemize}
%
%\begin{table}
%\centering
%\begin{tabular}{| l | r | r | r | r | r |}
%\hline
% & 2018 & 2018\footnote[frame]{meteorological conditions as in 2016} & Policy & 2030 & 2030 \\
% & [GW] & [TWh] & [TWh] & [TWh] & [GW] \\ \hline \hline
%Solar PV & 1.44 & 1.23 & +11 & 12.23 &  14.27 \\ \hline
%Wind (onshore) & 3.05 & 6.14 & +10 & 16.14 & 8 \\ \hline
%Hydro (run-of-river) & 5.72 & 28.34 & +5 & 33.34 & 6.73 \\ \hline
%Biomass & $\sim 1$ & 4.78 & +1 & 5.78 & $>2$ \\ \hline
%\end{tabular}
%\end{table}
%\end{frame}
%
%
%\begin{frame}
%\frametitle{Why wind is not solar}
%\framesubtitle{A side note on imperfect substitutes}
%%\begin{textblock*}{4cm}(0.75cm, 2.75cm)
%%Aggregate profile for Austria
%\begin{textblock*}{4cm}(0.75cm, 2.75cm)
%\begin{itemize}
%\item secure system operation requires $S=D$ at any point in time
%\item electricity not easily storable
%\item >>transforming<< electricity feed-in to end-use electricity is costly
%\end{itemize}
%\end{textblock*}
%
%\only<2>{
%\begin{textblock*}{10cm}(5.5cm, 2.25cm) % {block width} (coords)
%\centering
%Aggregate generation profile in Austria
%\includegraphics[width=\textwidth, keepaspectratio]{./figures/wind_not_pv}
%\end{textblock*}
%}
%\end{frame}
%

    \begin{frame}
        \frametitle{The case of Austria}
        \framesubtitle{Some Stylized Facts}
        \begin{itemize}
            \item Social conflict around the large-scale expansion of onshore wind power
            \item Solar PV is the only feasible large-scale substitute for wind power
            \begin{itemize}
                \item Use of nuclear power banned in a 1978 referendum
                \item Hydro power potential largely exploited
                \item Biomass neither ecologically nor economically sustainable
            \end{itemize}
            \item Germany is Austria's most important electricity trading partner, by far
        \end{itemize}
    \end{frame}
%
%\begin{frame}
%\frametitle{The problem from a social planner's perspective}
%How to account for local negative externalities in renewables expansion planning?
%
%\bigskip
%\textcolor{bokugreen}{The cost of undisturbed landscapes}
%\begin{itemize}
%\item assess energy system effects and costs of substituting wind power with solar PV
%\end{itemize}
%
%\bigskip
%\textcolor{bokugreen}{The value of undisturbed landscapes}
%\begin{itemize}
%\item estimates of the negative external effect of wind turbines reported in literature
%\end{itemize}
%\end{frame}

    \begin{frame}
        \frametitle{The case of Austria}
        \framesubtitle{Modeling the Austrian Power System}
        \begin{itemize}
            \item Resemble prospective power system in Austria in 2030
            \item Set policy target of meeting at least $90\%$ of demand in 2030 from domestic renewable sources
            \medskip
            \item Include most important electricity trading partner Germany
            \item Incorporate announced electricity system targets for Germany in 2030
            \begin{itemize}
                \item nuclear phase-out
                \item partial coal exit
                \item expansion of renewable capacities in line with EEG 2017
            \end{itemize}
            \medskip
            \item Simulate operation of and investment in prospective electricity system with \emph{medea}
        \end{itemize}
\end{frame}

\begin{frame}
\frametitle{Power system model \emph{medea}}
%\framesubtitle{Marketing slide}
\begin{footnotesize}
\begin{columns}[t]
\begin{column}{0.325\textwidth}
\textbf{Objective}
\begin{itemize}
\item minimize total system cost
	\begin{itemize}
	\begin{footnotesize}
	\item fuel and \ce{CO2} cost
	\item O\&M cost
	\item capital cost
	\end{footnotesize}
	\end{itemize}
\end{itemize}
\end{column}

\begin{column}{0.32\textwidth}  %%<--- here
\textbf{Decision variables}
\begin{itemize}
	\item hourly dispatch
	\item inter-zonal electricity trade
	\item investment in power plants, storages, and transmission
\end{itemize}
\end{column}

\begin{column}{0.32\textwidth}  %%<--- here
\textbf{Constraints}
\begin{itemize}
\item market clearing
\item capacity constraints
\item co-generation \& fuel use
%\item storage volume
\item system service requirement
\item inter-zonal electricity trade
\end{itemize}
\end{column}
\end{columns}
\end{footnotesize}
\bigskip
\begin{footnotesize}
\begin{columns}[t]
\begin{column}{0.325\textwidth}
\textbf{Economic assumptions}
\begin{itemize}
\item perfect competition
\item perfect foresight
\item price-inelastic demand
\end{itemize}
\end{column}

\begin{column}{0.325\textwidth}
\textbf{Resolution}
\begin{itemize}
\item hours (one year)
\item bidding zones
\item 41 technologies
\end{itemize}
\end{column}

\begin{column}{0.325\textwidth}
\textbf{Implementation}
\begin{itemize}
\item linear program
\item python \& GAMS
\end{itemize}
\end{column}
\end{columns}
\end{footnotesize}
\bigskip

\medskip
\begin{center}
    \emph{medea} is available on \textcolor{blue}{\href{https://github.com/inwe-boku/medea}{github.com/inwe-boku/medea}}
    under an open MIT license
\end{center}
\end{frame}
\note{
medea is a partial-equilibrium model of the coupled electricity and district heating sectors.
Operation of and investment in conventional and renewable power plants, storages and transmission lines are determined simultaneously for one year.
Results include hourly dispatch, investment, inter-zonal trade, electricity and heat prices, profits, and emissions of air pollutants.
The model accounts for various technical characteristics of power systems, such as co-generation of heat and power, balancing requirements, 
}

\section{Results}

\begin{frame}
\frametitle{The cost of undisturbed landscapes}

\begin{textblock*}{10cm}(0.75cm, 1.5cm) % {block width} (coords)
\includegraphics[width=\textwidth, keepaspectratio]{./figures/undisturbed_base}
\end{textblock*}

\begin{textblock*}{4.5cm}(10.9cm, 1.5cm)
\begin{small}
\begin{table}
\begin{tabular}{p{1.8cm} | r l}
Capital cost of solar PV & $630$ & \SI{}{\text{\euro}\per\kilo\watt\text{p}}
\end{tabular}
\end{table}
\smallskip
\quad $\rightarrow$ about $\sfrac{2}{3}$ rooftop PV, \\
\quad $\sfrac{1}{3}$ open space PV

\vspace{4.25cm}
\quad wind power and solar PV \\ \quad are substitutes
\end{small}
\end{textblock*}

\begin{textblock*}{9.5cm}(1.5cm, 8.2cm)
\begin{tikzpicture}%[thick]
    \draw (5.75,0) -- (9,0);
    \draw[{Latex[length=2mm]}-] (0,0) -- (3.25,0);
    \draw (4.5,0.0) node[] {\scriptsize more wind power};
\end{tikzpicture}

\begin{tikzpicture}%[thick]
    \draw[-{Latex[length=2mm]}] (5.5,0) -- (9,0);
    \draw (0,0) -- (3.5,0);
    \draw (4.5,0.0) node[] {\scriptsize more solar PV};
\end{tikzpicture}
\end{textblock*}

\end{frame}

\begin{frame}
\frametitle{The cost of undisturbed landscapes}
\framesubtitle{Cost with restricted wind power}
\begin{textblock*}{10cm}(0.75cm, 2.0cm) % {block width} (coords)
\includegraphics[width=\textwidth, keepaspectratio]{./figures/cost}
\end{textblock*}

\begin{textblock*}{4cm}(11cm, 2.25cm)
\begin{small}
    \begin{table}
        \begin{tabular}{p{1.8cm} | p{0.3cm} l}
            Capital cost of solar PV & $630$ & \SI{}{\text{\euro}\per\kilo\watt\text{p}} \\
            \ce{CO2} price & $90$ & \SI{}{\text{\euro}\per\mega\watt\hour}
        \end{tabular}
    \end{table}
\end{small}
\end{textblock*}

\end{frame}

\begin{frame}
\frametitle{The cost of undisturbed landscapes}
\framesubtitle{System operation with restricted wind power}
\begin{textblock*}{10cm}(0.75cm, 2.0cm) % {block width} (coords)
\includegraphics[width=\textwidth, keepaspectratio]{./figures/sysops}
\end{textblock*}

\begin{textblock*}{4cm}(11cm, 2.25cm)
\begin{small}
    \begin{table}
        \begin{tabular}{p{1.8cm} | p{0.3cm} l}
            Capital cost of solar PV & $630$ & \SI{}{\text{\euro}\per\kilo\watt\text{p}} \\
            \ce{CO2} price & $90$ & \SI{}{\text{\euro}\per\mega\watt\hour}
        \end{tabular}
    \end{table}
\end{small}
\end{textblock*}

\end{frame}

\section{Sensitivity}
\begin{frame}
\frametitle{The cost of undisturbed landscapes}
\framesubtitle{Sensitivity to capital cost of solar PV}
\begin{textblock*}{10cm}(0.75cm, 2.0cm) % {block width} (coords)
\includegraphics[width=\textwidth, keepaspectratio]{./figures/sens_nrg}
\end{textblock*}

\begin{textblock*}{4cm}(11cm, 2.25cm)
\begin{small}
    Capital cost estimates for solar PV in 2030
    \begin{table}
        \begin{tabular}{p{1.8cm} | p{0.3cm} l}
            Small-scale rooftop & $830$ & \SI{}{\text{\euro}\per\kilo\watt\text{p}} \\
            Utility-scale open space & $280$ & \SI{}{\text{\euro}\per\kilo\watt\text{p}} \\
        \end{tabular}
    \end{table}
\end{small}
\end{textblock*}
\end{frame}

\begin{frame}
\frametitle{The cost of undisturbed landscapes}
\framesubtitle{Sensitivity to capital cost of solar PV}
\begin{textblock*}{10cm}(0.75cm, 2.0cm) % {block width} (coords)
\includegraphics[width=\textwidth, keepaspectratio]{./figures/undisturbed_low}
\end{textblock*}

\begin{textblock*}{4cm}(11cm, 2.25cm)
\begin{small}
    \begin{table}
        \begin{tabular}{p{1.8cm} | p{0.3cm} l}
            Capital cost of solar PV & $280$ & \SI{}{\text{\euro}\per\kilo\watt\text{p}}
        \end{tabular}
    \end{table}
\end{small}
\end{textblock*}
\end{frame}

\section{Discussion \& Conclusion}
\begin{frame}
    \frametitle{Discussion of results}
    \begin{itemize}
        \item Renewable resource quality held constant as capacity is expanded
        \item Sub-national electricity transmission and distribution grids neglected
        \item Technical operation of generators not fully represented \\(e.g. no unit commit, simplified balancing)
        \item Electricity market-splitting has increased market concentration
        \item Announced policy necessarily turns Austria into a net exporter of electricity\\
        $\rightarrow$ "loop-flows" potentially avoided \\
        $\rightarrow$ artificial transmission restriction between DE and AT could be eliminated
%\item use Austrian landscape for "overinvesting" in wind and exporting to Germany?
    \end{itemize}
\end{frame}

\begin{frame}
    \frametitle{Conclusions}
    \begin{itemize}
        \item We demonstrated that restricting wind power can come at significant opportunity cost
        \item Socially optimal wind power deployment depends on cost and benefits of undisturbed landscapes
        \item Our analysis constitutes one of two pillars needed to determine a spatially explicit, socially
        optimal renewables expansion
        \begin{itemize}
            \item[$\rightarrow$] need to complement findings with a spatially resolved assessment of the negative
            externalities of wind turbines, for example approximated by wind turbine impact on property prices
        \end{itemize}
        \medskip
        \item For the case of Austria, we find
        \begin{itemize}
            \item little social value lost from substituting onshore wind power with utility-scale, open-space solar PV,
            \emph{if} \ce{CO2} emissions are valued at \SI{30}{\text{\euro}\per\mega\watt\hour} or lower
            \item potential gains from wind power \emph{if} (i) there is a preference for rooftop solar PV or (ii)
            the cost of \ce{CO2} emissions exceeds \SI{30}{\text{\euro}\per\mega\watt\hour}
            \item Gains from wind power deployment could be used to compensate individuals affected by local wind turbine
            externalities
        \end{itemize}
    \end{itemize}
\end{frame}

 %%%Last Frame
 {
    \usebackgroundtemplate{
        \begin{picture}(320,290)
        \hspace{9.5cm}
          \includegraphics[width=0.6\linewidth]{./logos/refuel_logo_with_text}
%        \begin{picture}(200,305)
%        \hspace{7.7cm}
%          \includegraphics[width=0.6\linewidth]{refuel_logo_with_text}
        \end{picture}
    }
    \begin{frame}
        \frametitle{Thank you!}
        \vspace{0.5cm}
        \url{https://refuel.world}\\
        \url{https://github.com/inwe-boku/medea}\\
        \url{sebastian.wehrle@boku.ac.at}

		\vspace{0.6cm}
        \qrcode[height=2cm]{http://github.com/inwe-boku/medea}\\

        \vspace{1.65cm}
        \includegraphics[height=0.7cm]{./logos/creative-commons-by}

        \smallskip
        \tiny{We gratefully acknowledge support from the European Research Council \\(``reFUEL'' ERC-2017-STG 758149).}

    \end{frame}
}   
%\appendix
%\backupbegin
%
%
%\begin{frame}
%\frametitle{The cost of undisturbed landscapes}
%\framesubtitle{Unconstrained deployment of wind and solar power}
%
%\begin{textblock*}{10cm}(0.75cm, 1.8cm) % {block width} (coords)
%\begin{table}
%\begin{tabular}{l r l}
%\emph{Results for Austria} \\ \hline \hline
%Added Solar PV & 0.0 & \SI{}{\tera\watt\hour} \\ \hline
%Added Wind & 31.0 & \SI{}{\tera\watt\hour} \\ \hline \hline
%System Cost\footnote[frame]{excluding cost for domestic transmission and distribution grids} & $3095.2$ & \SI{}{\mega\text{\euro}} \\ \hline
%Trade balance & $746.8$ & \SI{}{\mega\text{\euro}} \\ \hline
%Cost of air pollution & $184.4$ & \SI{}{\mega\text{\euro}} \\ \hline
%\ce{CO2} emissions & $5227.3$ & \SI{}{\kilo\tonne} \\ \hline
%Electricity generation & $92.8$ & \SI{}{\tera\watt\hour} \\ \hline
%Curtailment & $2.4$ & \SI{}{\tera\watt\hour} \\ \hline
%Net exports & $5.9$ & \SI{}{\tera\watt\hour} \\ \hline
%Wholesale electricity price & $56.05$ & \SI{}{\text{\euro}\per\mega\watt\hour} \\ \hline
%\end{tabular}
%%\caption{Results for Austria}
%\end{table}
%\end{textblock*}
%
%\begin{textblock*}{4cm}(11cm, 2.05cm)
%\begin{small}
%\begin{table}
%\begin{tabular}{p{1.8cm} | p{0.3cm} l}
%Capital cost of solar PV & $630$ & \SI{}{\text{\euro}\per\kilo\watt\text{p}} \\
%\ce{CO2} price & $90$ & \SI{}{\text{\euro}\per\mega\watt\hour}
%\end{tabular}
%\end{table}
%\end{small}
%\end{textblock*}
%
%\end{frame}
%
%
%\begin{frame}
%\frametitle{The value of undisturbed landscapes}
%\framesubtitle{Local Negative Externalities of Wind Turbines}
%%\begin{table}[h!]
%%\caption{Estimates of Wind Turbine Impact on Property Prices}
%\begin{table}
%\centering
%\begin{scriptsize}
%	\begin{tabular}{p{3.2cm} l l l p{1.1cm} p{2.4cm}}
%	\hline
%	Authors & Country & Sample & Framework\footnote[frame]{DID $\ldots$ difference-in-differences, SAR $\ldots$ spatial autoregressive, SFE $\ldots$ spatial fixed effects} & Impact Proxy\footnote[frame]{d $\ldots$ distance, v $\ldots$ visibility} & Estimated Impact \\
%	\hline \hline
%%	\cite{Sims2008} & Cornwall, UK & 2000 - 2007 & hedonic & visibility dummies & insignificant \\ \hline
%	Heintzelmann and Tuttle, 2012 & USA, NY & 2000 - 2009 & hedonic, SFE & d & $-3.6\%$ to $-9.6\%$, insignificant \\ \hline
%	Jensen et al., 2014 & DNK & 2000 - 2011 & hedonic, SAR & d, v & $-4.35\%$ \\ \hline
%	Lang et al., 2014 & USA, RI & 2000 - 2013 & DID & d, v & insignificant  \\ \hline
%	Gibbons, 2015 & GBR, EAW & 2000 - 2011 & DID & d, v & $-6\%$ \\ \hline
%	Hoen et al., 2015 & USA & 1996 - 2011 & DID & d & insignificant \\ \hline
%	Dröes and Koster, 2016 & NED & 1985 - 2011 & DID & d & $-1.4\%$\\ \hline
%	Sunak and Madlener, 2016 & GER, NRW & 1992 - 2010 & DID & v & $-9\%$ to $-14\%$ \\ \hline
%	Jensen et al., 2018 & DNK & - & hedonic, SAR & d, \# of turbines & $-0.2\%$ to $-1.1\%$ \\ \hline
%	Kussel et al., 2019 & GER & 2007 - 2015 & DID & d & $-6.0\%$ \\ \hline \hline
%	\end{tabular}
%%\label{tab:negex}
%\end{scriptsize}
%\end{table}
%\end{frame}
%
%
%\begin{frame}
%\frametitle{Compensating for local negative externalities}
%\framesubtitle{An example}
%\begin{textblock*}{7cm}(0.75cm, 2.0cm) % {block width} (coords)
%OC of wind turbines in Austria
%\vspace{-0.41cm}
%\begin{table}
%%\centering
%\begin{footnotesize}
%	\begin{tabular}{c | c | c | c}
%	$oc_w$ & $\delta$ & NPV $oc_w$ & NPV $oc_w$ \\
%	$[\text{\euro}\, \text{MW}^{-1} \text{a}^{-1}]$ & & $[\text{\euro}\, \text{MW}^{-1}]$ & $[\text{\euro} \, \text{turbine}^{-1}]$ \\ \hline \hline
%	$40\,000$ & $0.05$ & $563\,758$ & $1\,973\,152$ \\ \hline
%	$40\,000$ & $0.03$ & $696\,526$ & $2\,437\,841$ \\ \hline
%	$10\,000$ & $0.05$ & $140\,939$ & $493\,288$ \\ \hline
%	$10\,000$ & $0.03$ & $171\,131$ & $609\,460$ \\ \hline
%	\end{tabular}
%\end{footnotesize}
%\end{table}
%\end{textblock*}
%
%\begin{textblock*}{7cm}(8.25cm, 2.0cm) % {block width} (coords)
%Property value loss in Germany
%%\vspace{-0.5cm}
%\begin{table}
%%\centering
%\begin{footnotesize}
%	\begin{tabular}{c | c | c | c}
%	  & value & impact & loss \\
%	  & \euro & \% & \euro \\ \hline \hline
%	mean & $273\,786$ & $0.06$ & $16\,427$ \\ \hline
%	mean & $273\,786$ & $0.10$ & $27\,379$ \\ \hline
%	mean + 1 s.d. & $476\,923$ & $0.06$ & $28\,615$ \\ \hline
%	mean + 1 s.d. & $476\,923$ & $0.10$ & $47\,692$ \\ \hline
%	\end{tabular}
%\end{footnotesize}
%\end{table}
%\end{textblock*}
%
%\begin{textblock*}{14.5cm}(0.75cm, 6.0cm) % {block width} (coords)
%\centering
%Compensation potential
%\vspace{-0.1cm}
%\begin{table}
%\begin{footnotesize}
%	\begin{tabular}{c | c | c | c | c}
%	$oc_w$  & $\delta$ & value & impact & \# property owners \\ \hline \hline
%	$40\,000$ & $0.05$ & $273\,786$ & $0.06$ & $120$ \\ \hline
%	$40\,000$ & $0.05$ & $273\,786$ & $0.10$ & $72$ \\ \hline
%	$40\,000$ & $0.03$ & $476\,923$ & $0.06$ & $69$ \\ \hline
%	$40\,000$ & $0.03$ & $476\,923$ & $0.10$ & $41$ \\ \hline
%%	$10\,000$ & $0.05$ & $476\,923$ & $0.06$ & $1$ \\ \hline
%	$10\,000$ & $0.03$ & $476\,923$ & $0.10$ & $18$ \\ \hline
%	\end{tabular}
%\end{footnotesize}
%\end{table}
%\end{textblock*}
%
%\end{frame}
%\backupend
\end{document}